Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-27 LME-MPM.tex

\usepackage{lineno}
% \usepackage{hredyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies. \PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

@@

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute \PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing \PNA{somehow} the accuracy. \MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several challenging test cases \PNA{on the elastic regime}: the
propagation of shock waves in a elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method is composed by three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition.

\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied. \PNA{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected. \PNA{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

In order to obtain a finite set of equations, in contrast with the
\acrshort{fem}, in \acrshort{mpm} a double discretization procedure is
performed. First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where Riemann \PNA{ALGUNA CITA AQUI}
integral definition is recovered as an addition of a finite set of points, and
their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Following, the aforementioned second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes. \PNA{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Mesh-particles-back}}
  \caption{Description of the spatial discretization for domain presented in the
    figure~\ref{fig:Continuum-solid}. Blue mesh represent the
    background computational support, and the red mesh conforms the
    discretized continuum body.}
  \label{fig:MPM-discretization}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},  \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
approximating the displacement field of the particle
$p$ as $\vec{u}_{p} = N_{Ip} \vect{u}_I$, $\vec{u}^{\psi}_{p} = N_{Ip} \vect{u}_I^{\psi}$,
and its gradient as $\tens{\varepsilon}_p = (\vect{u}_I \otimes
\Grad{N_{Ip}})^s$, $\tens{\varepsilon}_p^{\psi}= (\vect{u}_I^{\psi} \otimes \Grad{N_{Ip}})^s$.
nodal balance of forces of the continuum yields:
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$ and $\tens{m}_{IJ}$, the nodal mass matrix, is obtained through:
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} = N_{Ip} m_p N_{Jp}.
\end{equation}
In order to improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Following, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = N_{Ip}\ \vec{b}_{p}\ m_p  + N_{Ip}\ \vec{t}^s_{p}\ m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor rate, $\dot{ \tens{\varepsilon}}_{p}$, as a measure of the time derivative of the strain tensor, is updated employing the velocity at the background mesh by the equation:
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, mass conservation is guaranteed by enforcing the null value of
the material derivative of the density field $\frac{D \rho}{D t} = 0$.
This leads to a suitable equation to update the density field:
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{trace} \left( \dot{\tens{\varepsilon}} \right).
\end{equation}
Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order \PNA{time} integration scheme is required. Therefore, time is
discretized into a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation is solved, the values at the nodes are
interpolated back into the particles, which are advected
to the new position through:
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
\PNA{Traditionally, Eqs.
\eqref{eq:particle_balance_forces3} and ~\eqref{eq:Updated_Lagrangian},
are solved with an explicit forward Euler algorithm. In the following subsection, this and the proposed schemes are described}.

@@

\subsection{Local \textit{Max-Ent} approximants}
\label{sec:local-max-ent}
The popularity of the \acrshort{mpm} has increased notoriously during
the recent years due to its ability to deal with large strain problems
without mesh distorsion issues inherent to mesh based methods like
\acrshort{fem}, see Wi{\c{e}}ckowski \cite{Wieckowski2004}. However, in the simulations
made with the original \acrshort{mpm}, numerical noises appear when particles
cross the cell boundaries. \PNA{Solving this issue is the main goal of the employment of the \acrshort{LME} shape functions}.

>>>>

\usepackage{lineno}
\usepackage{hyperref}
\usepackage{float}
\usepackage{subfig}
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

\section{Introduction}
\label{intro}
Since the proposal of \acrshort{mpm} by Sulsky {\it  et al.}
(1994)~\cite{Sulsky1994} as a generalization to solids of the \acrfull{flip} method~\cite{Brackbill1986}, its popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.\PNA{One of the main fields where this method is strong is the dynamic one since the original time integration scheme proposed was the \acrfull{fe}~\cite{Sulsky1994}, carried out in a explicit manner. However, in this type or problems, the main instabilities of the original \acrshort{mpm} are even marked.}

@@

This document adopts the \acrfull{lme}, or Local \textit{Max-Ent} approximates, as a robust substitute\PNA{of the aforementioned} shape functions in \acrshort{mpm}. First introduced by Arroyo \& Ortiz
(2006)~\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
\acrshort{fem} and meshfree interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The \acrshort{lme} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. \acrshort{lme} basis functions possess many desirable properties for
meshfree algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for related functions~\cite{Arroyo2006}. Important contributions on the Maximum-Entropy have been made by Sukumar and coworkers~\cite{Sukumar15} with Cell-based techniques and the ones carried within the \acrfull{otm} method. The latter methodology has been proven to have a good performance under
the dynamic regime by other researchers, being important the contributions of Li {\it et al.} (2012)~\cite{Li2012} and Navas {\it et al.}
(2018)~\cite{Navas:17b,Navas2018a} in the explicit regime and Navas {\it et al.}~\cite{Navas2016,Navas2016b,Navas:17c} and Wriggers and coworkers~\cite{Wriggers18} with implicit schemes. And more
recently, under \acrshort{mpm} framework, the work made by Wobbes {\it et
  al.}(2020)~\cite{Wobbes2020}.\PNA{The proposed research delves into the benefits of
the regularization parameter, \gls{LME-beta}, and the analogy of the different obtained shape functions derive by the tuning of this parameter and the traditional \acrshort{mpm} ones.}

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretization technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers has proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing\PNA{somehow} the accuracy.\MMP{One of the first attempts of it was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} to downsize this spurious noises add a nonvoiscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. Within the explicit
time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort.} In this paper a less time consuming and high efficient explicit predictor-corrector integration method has been
proposed. It consists in an accommodation of the traditional \acrsull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and its good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

The aim of this document is to mitigate the spurious oscillations due
inaccuracies in both space and time discretization by the employment of a
suitable combination of the \acrshort{lme} family shape functions, and the proposal of an explicit predictor-corrector scheme. The advantages of
this approach will be illustrated through several challenging test cases\PNA{on the elastic regime}: the
propagation of shock waves in a elastic bar and the response of a block of soil gradually loaded with gravitational forces.

@@

The aim of this section is to provide an overview of the standard
explicit \acrshort{mpm} algorithm~\cite{Sulsky1994}. Without loosing
generality, the method is composed by three main steps: (i) a
variational recovery process, where particle data is projected to the
grid nodes, (ii) an Eulerian step, where balance of momentum equation
is expressed as a nodal equilibrium equation through a \acrshort{fem}-like
procedure, and finally (iii) a Lagrangian advection of the
particles. In consequence, \acrshort{mpm} can be regarded as a
Lagrangian-Eulerian method where particles carry on all the physical
information and a set of background nodes is employed to
compute the equilibrium equation. In what follows, we will adopt the
following convention. Three kind of subscript or superscript are used
within paper. The subscript $\Box_p$ is used to define a particle
variable. While the subscript $\Box_I$ is reserved in this notation for denoting nodal
variables. And finally, the superscript $\Box^{\psi}$ involves a
virtual magnitude. For the operators, the convention is : $\dot{\Box}$ and
$\ddot{\Box}$ for the first and second time derivative, $\qued \otimes
\qued$ means the dyadic operator, $\Box \cdot \Box$ and $\Box \colon \Box$ means the
first and second contraction of a tensor, $\Div{\Box}$ denotes the
divergence operator, and finally $\Grad{\Box}$ and $\GradS{\Box}$
denotes the gradient and its symmetric part. Einstein subscripts
convention is adopted therefore repeated index means addition.\PNA{Following, the \acrshort{mpm} methodology, the explicit predictor-corrector scheme and \acrshort{lme} approximation shape functions are describe in subsection \ref{sec:derivation-mpm},\ref{sec:epc-algor-mpm} and \ref{sec:local-max-ent} respectively.}

@@

In \acrshort{mpm} the continuum mechanics approach is considered. So on, let
define a continuum \gls{domain}  occupied by an elastic body like
the sketched in the figure~\ref{fig:Continuum-solid}, and \gls{contour} the boundaries of the domain defined by $\partial \varOmega
= \Gamma_d \bigcup \Gamma_n$ and $\Gamma_d \cap  \Gamma_n = \emptyset$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Continuum-solid}}
  \caption{Description of the boundary-value-problem in a
    continuum. Red lines represents the closure $\partial \varOmega$
    of the domain $\varOmega$ represented in gray.}
  \label{fig:Continuum-solid}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this context the field \gls{u} allows to describe the \textit{global state}
of the system. Now the variable $\phi =
(\tens{\varepsilon},\tens{\sigma})$ is defined as the set of \textit{local
  states} at any point of the continuum which can be derived from the
field \gls{u} through the following set of governing equations and
restrictions that must be satisfied.\PNA{First (i) the \textit{compatibility
  equation}, where the the strain field \gls{strain} is extracted from \gls{u} is defined as:}
\begin{equation}
  \label{eq:Compatibility-equation}
  \tens{\varepsilon} = \GradS{\vect{u}},
\end{equation}
together with essential boundary conditions of Dirichlet type \gls{dirichlet-boundary}. An additional consideration over the strain field is
the assumption of inifinitesimal strain, therefore second order terms
in the spatial derivatives can be neglected.\PNA{Therefore, the stress field \gls{stress} will be considered the corresponding conjugate variable for the strain field, being the one} which satisfies (ii) the \textit{conservation of
  momentum equation}:
\begin{equation}
  \label{eq:Balance-momentum}
\rho \frac{D\vect{v}}{Dt} = \Div{\tens{\sigma}} + \rho \vect{b}
\end{equation}
together with the natural boundary conditions of the Neumann type \gls{neumann-boundary}.  An aditional component will be (iii) the constitutive equation as a linear
application from $\Re^n$ to $\Re^n$, which relates the
strain tensor with the stress tensor:
\begin{equation}
  \label{eq:Constitutive-equation}
\tens{\sigma} = \tens{D} \colon \tens{\varepsilon}.
\end{equation}
\PNA{In this research, plane strain Linear Elasticity has been considered. Thus the constitutive tensor, \tens{D}, is the well known linear elastic one.} The final restriction is (iv) the mass conservation, which can be obtained by
setting to zero the total derivative of the density field,
\begin{equation}
  \label{eq:Rho-material-derivative}
  \frac{D \rho}{D t} = \dot{\rho} + \rho \Div{\vect{v}} = 0.
\end{equation}

@@

In order to obtain a finite set of equations, in contrast with the
\acrshort{fem}, in \acrshort{mpm} a double discretization procedure is
performed. First, the continuum \gls{domain} is discretized with a finite sum of material points (also denominated particles the manuscript). Each material point
represents a part of the discretized domain $\varOmega_p \subset
\varOmega$ with $p = 1,2\ldots ,N_p$ where $N_p$ is the number of
particles. The material point location, $\vec{x}_p$, is defined at the centroid
of each $\Omega_p$ (see figure~\ref{fig:MPM-discretization} for details).
Initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point, which also owns the
virtual displacement field $\vect{u}^{\psi}_{p}$. Therefore, employing
the definition of the material integral, where Riemann
integral definition \cite{Riemann_1854} is recovered as an addition of
a finite set of points, and their volumes are interpreted as quadrature weights. Consequently,
individual terms in \eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \frac{d\vec{v}}{dt} \cdot \vect{u}^{\psi}} =
    \frac{d\vec{v}_{p}}{dt} \cdot \vect{u}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{u}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{u}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{u}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{u}^{\psi}} = \vec{t}^s_{p} \cdot \vect{u}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Following, the aforementioned second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated, being $N_n$ the number of grid
nodes.\PNA{Spatial derivatives, such as gradients and
divergences, are computed through the support of the background mesh.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.5\hsize}{!}{
    \includegraphics[width=\textwidth]{Figures/Mesh-particles-back}}
  \caption{Description of the spatial discretization for domain presented in the
    figure~\ref{fig:Continuum-solid}. Blue mesh represent the
    background computational support, and the red mesh conforms the
    discretized continuum body.}
  \label{fig:MPM-discretization}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces},  \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
approximating the displacement field of the particle
$p$ as $\vec{u}_{p} = N_{Ip} \vect{u}_I$, $\vec{u}^{\psi}_{p} = N_{Ip} \vect{u}_I^{\psi}$,
and its gradient as $\tens{\varepsilon}_p = (\vect{u}_I \otimes
\Grad{N_{Ip}})^s$, $\tens{\varepsilon}_p^{\psi}= (\vect{u}_I^{\psi} \otimes \Grad{N_{Ip}})^s$.
nodal balance of forces of the continuum yields:
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$ and $\tens{m}_{IJ}$, the nodal mass matrix, is obtained through:
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} = N_{Ip} m_p N_{Jp}.
\end{equation}
In order to improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Following, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = N_{Ip}\ \vec{b}_{p}\ m_p  + N_{Ip}\ \vec{t}^s_{p}\ m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor rate, $\dot{ \tens{\varepsilon}}_{p}$, as a measure of the time derivative of the strain tensor, is updated employing the velocity at the background mesh by the equation:
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, mass conservation is guaranteed by enforcing the null value of
the material derivative of the density field $\frac{D \rho}{D t} = 0$.
This leads to a suitable equation to update the density field:
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{trace} \left( \dot{\tens{\varepsilon}} \right).
\end{equation}
Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order\PNA{time} integration scheme is required. Therefore, time is
discretized into a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation is solved, the values at the nodes are
interpolated back into the particles, which are advected
to the new position through:
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
\PNA{Traditionally, Eqs.
\eqref{eq:particle_balance_forces3} and ~\eqref{eq:Updated_Lagrangian},
are solved with an explicit forward Euler algorithm. In the following subsection, this and the proposed schemes are described}.

@@

\subsection{Local \textit{Max-Ent} approximants}
\label{sec:local-max-ent}
The popularity of the \acrshort{mpm} has increased notoriously during
the recent years due to its ability to deal with large strain problems
without mesh distorsion issues inherent to mesh based methods like
\acrshort{fem}, see Wi{\c{e}}ckowski \cite{Wieckowski2004}. However, in the simulations
made with the original \acrshort{mpm}, numerical noises appear when particles
cross the cell boundaries.\PNA{Solving this issue is the main goal of the employment of the \acrshort{LME} shape functions}.
